---
title: "Lumpkin Foundation - Illinois Bundleflower Final Report"
author: "Spencer Barriball, Bo Meyering, Brandon Schlautman"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library import, echo=FALSE, results='hide', include=FALSE}
library(tidyverse, warn.conflicts=FALSE, quietly=TRUE)
library(lme4, warn.conflicts=FALSE, quietly=TRUE)
library(sommer, warn.conflicts=FALSE, quietly=TRUE)
library(emmeans, warn.conflicts = FALSE, quietly = TRUE)
library(data.table, warn.conflicts=FALSE, quietly=TRUE)
library(knitr)
library(kableExtra)
library(multcomp)
library(lmerTest)
```

```{r Data Import and Cleaning, include=FALSE, echo=FALSE}
# Read in trial layout and fix sample_name
trial_layout <- fread('../data/trial_layout_metadata.csv') %>% 
        mutate(sample_name = paste('LMP', plots, sep='-')) %>% 
        filter(!trts %in% c('HI-GEST 660', 'Kura Clover', 'Sainfoin')) %>% 
        dplyr::select(-V1) %>% 
        rename("sample_id" = "trts")

# Grab only unique sample_ids
sample_ids <- data.frame(sample_id = unique(trial_layout$sample_id))
# Read in germination data and rearrange
germination_data <- fread('../data/lumpkin_ibf_germination_test.csv') %>% 
        mutate(sample_id = factor(sample_id)) %>% 
        arrange(desc(germination_pct), hard_pct) %>% 
        right_join(sample_ids)

# Read in forage yield data, join to trial layout
yield_data <- fread('../data/lumpkin_ibf_forage_yield.csv') %>% 
        left_join(trial_layout, by='sample_name') %>% 
        dplyr::select(sample_name, sample_id, block, rep, row, column, forage_yield_kg_ha)

# Read in the wet chemistry results
wetchem <- fread('../data/lumpkin_ibf_wet_chemistry_results.csv')

# Pivot wetchem longer
wetchem_longer <- wetchem %>% 
        dplyr::select(-c(moisture_pct, dry_matter_pct)) %>% 
        pivot_longer(cols = c(crude_protein_pct:nel_adf), names_to='test', values_to='results')

# Summarize the yield data
yield_summaries <- yield_data %>% 
        group_by(sample_id) %>% 
        summarize(mean_yield = mean(forage_yield_kg_ha, na.rm=TRUE),
                  sd_yield = sd(forage_yield_kg_ha, na.rm=TRUE),
                  n = n(),
                  se_yield = sd_yield / sqrt(n))

# Pivot germination_data longer
germination_long <- germination_data %>% 
        pivot_longer(cols = c(germination_pct, hard_pct), names_to = 'key', values_to='values')

# Read in the model outputs from mlflow runs
model_results = fread("../model_output/runs.csv") %>% 
        mutate(`Test RMSE` = round(`Test RMSE`, 3),
               `Train RMSE` = round(`Train RMSE`, 3)) %>% 
        arrange(`Forage Trait`, desc(`Forage Size`)) %>% 
        dplyr::select(`Forage Trait`, `Forage Size`, `Num. Components`, `Train RMSE`, `Test RMSE`)




```

# Introduction and Summary of Work

# Materials and Methods

## Accession Germination Test

## Field Trial Design

## Harvesting

## Sample Processing

## NIR Spectra Sampling
Plant tissues from harvet were ground and filtered using a 1mm sieve .... and scanned using the NeoSpectra NIR handheld scanner ... FILL THIS OUT

## Wet Chemistry
A total of 75 samples out of the entire population were selected for targeted wet chemistry analysis. Samples were selected out of the entire population using the Kennard Stone algorithm to select samples with the greatest Euclidean distance from each other, ensuring that the samples reflected the availble NIR variance in the population. Samples were processed by xyz method, dried and sent to ABC labs for targeted wet chemistry analysis.. FILL THIS OUT


## NIR Modeling
A total of ten wet chemistry metrics were selected for modeling. These included forage moisture content (%), forage dry matter content (%), ADF (units), NDF (units), relative feed value (units), total digestible nutrients (units), net energy gain (Mcal/cwt), net energy maintenance (Mcal/cwt), net energy lactation (Mcal/cwt), and forage yield (kg/Ha) all calculated on a dry matter basis. Spectra were presented as the percent reflectance per wavelength measured. Low quality spectra, i.e. low variance among the response values for all wavelengths, were filtered out of the dataset programatically. All spectra were then linearly resampled by wavelength to ensure equal wavelength intervals between each response value. The spectra were smoothed and detrended using a Savitsky-Golay filter by fitting a $5^{th}$ order polynomial over a window length of 15 frames and taking the second derivative. Standard mean subtraction scaling was applied per wavelength, but were not scaled by the colun variance. Training and testing populations (70% and 30%, respectively) were selected from the calibration samples by using the Kennard-Stone selection algorithm. A partial least squares regression (PLSR) model was fit to the data for each of the target variables, allowing the number of model components to vary from 1 to 16 total components. Root Mean Squared Error (RMSE) defined below was used to evaluate each models performance on the training and testing populations. 
$$
RMSE = \sqrt{\frac{\sum^N_{i=1}(y_i - \hat{y_i})^2}{N}}
$$
where $y_i$ represents the ground truth target value of the $i^{th}$ sample and $\hat{y_i}$ represents the model prediction for that sample. The model with the lowest testing population RMSE was selected as the final model.

## Forage Quality Breeding BLUE and Repeatability
We modeled each of the forage quality trait as $ X $

Repeatabilty was then defined as the variance of the random effect 'Sample ID' divided by the total variance in the model 
$$
\frac{\sigma^2_G}{\sigma^2_G + \frac{\sigma^{2}_{resid.}}{r}}
$$ 
where $r$ is equal to the total number of replicates in the experiment, $\sigma^2_G$ is the sample variance from the random accession effects, and $\sigma^2_{resid.}$ is the residual model variance.

## Selection Index

We defined a selection index based on the importance of 3 traits: Forage yield, total digestible nutrients, and crude protein such that the selection index for IBF accesssion $i$ can be represented as:
$$
SI_i = w \cdot x_i
$$
where $w$ is a column vector of weights that sum to 1, and $x_i$ is a vector of traits from the $i^{th}$ accession expressed as $x_i = [\text{forage yield}, \text{tdn} , \text{crude protein}]$

# Results
## Germination Results

```{r Germination results plot, fig.width=10}
germination_plot <- ggplot(germination_long, aes(x=reorder(sample_id, -total_viable_pct), y=values, fill=key))+
        geom_bar(stat = 'identity', position='stack', width=.9)+
        theme_light()+
        scale_fill_viridis_d(option='D', begin=0.2, end=0.8)+
        labs(x = 'Accession', y='Total Viability', fill="Germination", title="Total Accession Seed Viability")+
        ylim(min=0, max=100)+
        theme(axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              panel.grid.major.x = element_blank(),
              panel.grid.minor.x = element_blank())

germination_plot
```



## Yield Data
```{r}
trial_layout <- fread('../data/trial_layout_metadata.csv') %>% 
        mutate(sample_name = paste('LMP', plots, sep='-')) %>% 
        dplyr::select(-V1)
yield_data <- fread('../data/lumpkin_ibf_forage_yield.csv') %>% 
        left_join(trial_layout, by='sample_name') %>% 
        dplyr::select(sample_name, trts, block, rep, row, column, forage_yield_kg_ha) %>% 
        rename('sample_id' = 'trts')

# Filter out plots with really low yield
yield_data <- yield_data %>% 
        filter(forage_yield_kg_ha > 500)

yield_summaries <- yield_data %>% 
        group_by(sample_id) %>% 
        summarize(mean_yield = mean(forage_yield_kg_ha, na.rm=TRUE),
                  sd_yield = sd(forage_yield_kg_ha, na.rm=TRUE),
                  n = n(),
                  se_yield = sd_yield / sqrt(n))

yield_mm <- lmer(forage_yield_kg_ha ~ sample_id + (1|block), data=yield_data, )
anova(yield_mm)
summary(yield_mm)
```


## NIR Modeling

```{r Model Results Import, echo=FALSE}
model_results = fread("../model_output/runs.csv") %>% 
        mutate(`Test RMSE` = round(`Test RMSE`, 3),
               `Train RMSE` = round(`Train RMSE`, 3))

kable(model_results, align = 'c')
```

```{r Yield boxplot, echo=FALSE, include=TRUE}
# Descending yield boxplot
median_values = yield_data %>% 
        group_by(sample_id) %>% 
        summarize(median_yield = median(forage_yield_kg_ha))

yield_data <- yield_data %>% 
        left_join(median_values)

yield_plot = ggplot(yield_data, aes(x=reorder(sample_id, -forage_yield_kg_ha, FUN=median), y=forage_yield_kg_ha, fill=median_yield))+
        geom_boxplot()+
        scale_fill_viridis_c(option='D')+
        geom_hline(yintercept=mean(yield_data$forage_yield_kg_ha), linewidth=1, linetype=4)+
        theme_light()+
        labs(x = 'Accession', y='Forage Yield (kg/Ha)', fill = 'Forage Yield')+
        theme(axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              panel.grid.major.x = element_blank(),
              panel.grid.minor.x = element_blank())

yield_plot
```

```{r}
# Descending yield boxplot
# median_values = ibf_10mm_summary %>% 
#         left_join(trial_layout %>% dplyr::select(c(sample_name, trts)), by='sample_name') %>% 
#         dplyr::select(-sample_name) %>% 
#         rename('sample_id' = 'trts') %>% 
#         group_by(sample_id) %>% 
#         summarize_all(median) %>% 
#         filter(!sample_id %in% c('HI-GEST 660', 'Kura Clover', 'Sainfoin'))
# 
# ibf_10mm_summary_df = ibf_10mm_summary %>% 
#         left_join(trial_layout %>% dplyr::select(c(sample_name, trts)), by='sample_name') %>% 
#         dplyr::select(-sample_name) %>% 
#         rename('sample_id' = 'trts')
# 
# 
# cp_plot = ggplot(
#         ibf_10mm_summary_df, 
#         aes(x=reorder(sample_id, -adf_pct, FUN=median), y=adf_pct)
#         )+
#         geom_boxplot()+
#         scale_fill_viridis_c(option='D')+
#         theme_light()+
#         labs(x = 'Accession', y='Forage Yield (kg/Ha)', fill = 'Forage Yield')+
#         theme(axis.text.x = element_blank(),
#               axis.ticks.x = element_blank(),
#               panel.grid.major.x = element_blank(),
#               panel.grid.minor.x = element_blank())
# 
# cp_plot
```


```{r}

ggplot(wetchem_longer, aes(x=results, group=test, fill=test))+
        geom_density()+
        facet_wrap(~test, scales = 'free', nrow=2)

crude_protein_plot = ggplot(wetchem, aes(x=reorder(sample_name, -crude_protein_pct, FUN=median), y=crude_protein_pct))+
        geom_boxplot()+
        scale_fill_viridis_c(option='D')+
        # geom_hline(yintercept=mean(yield_data$forage_yield_kg_ha), linewidth=1, linetype=4)+
        theme_light()+
        labs(x = 'Accession', y='Forage Yield (kg/Ha)', fill = 'Forage Yield')+
        theme(axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              panel.grid.major.x = element_blank(),
              panel.grid.minor.x = element_blank())

crude_protein_plot

```

```{r IBF prediction list and summary, echo=FALSE, warning=FALSE, message=FALSE}
ibf_predictions <- list("1mm" = NULL, 
                        "10mm" = NULL)

file_list = dir('../predictions')

for (file_name in file_list) {
        if (startsWith(file_name, 'ibf')) {
                # Extract the forage sample size
                forage_size = str_extract(file_name, "(?<=_)\\d+mm")
                
                # Extract the forage trait
                trait = str_extract(file_name, "(?<=ibf_).*?(?=_\\d+mm)")
                
                file_path = file.path('../predictions', file_name)
                predictions = fread(file_path)

                predictions <- predictions %>% 
                        rename(!!trait := 'value') # Rename the value column dynamically !!(name) evaluates name
                
                if (is.null(ibf_predictions[[forage_size]])) {
                        ibf_predictions[[forage_size]] = predictions
                } else {
                        ibf_predictions[[forage_size]] = ibf_predictions[[forage_size]] %>% 
                                left_join(predictions)
                }
                
                
        }        
}

ibf_1mm_summary <- ibf_predictions[['1mm']] %>% 
        group_by(sample_name) %>% 
        summarize_all(mean) %>% 
        dplyr::select(-rep) %>% 
        left_join(
                trial_layout %>% 
                        dplyr::select(c(sample_name, 1:3)),
                by=c('sample_name')
        ) %>% 
        rename('sample_id' = 'trts') %>% 
        dplyr::select(sample_id, block, plots, sample_name, everything())

ibf_10mm_summary <- ibf_predictions[['10mm']] %>% 
        group_by(sample_name) %>% 
        summarize_all(mean) %>% 
        dplyr::select(-rep) %>% 
        left_join(
                trial_layout %>% 
                        dplyr::select(c(sample_name, 1:3)),
                by=c('sample_name')
        ) %>% 
        rename('sample_id' = 'trts') %>% 
        inner_join(yield_data %>% dplyr::select(sample_name, forage_yield_kg_ha), by='sample_name') %>% 
        dplyr::select(sample_id, block, plots, sample_name, everything())

write_csv(ibf_1mm_summary, 'ibf_1mm_predictions.csv')
write_csv(ibf_10mm_summary, 'ibf_10mm_predictions.csv')


```

## Trait Predictions and Repeatability

```{r repeatability modeling}
# Function to calculate the repeatability of a trait
calc_repeatability <- function(model, r) {
        # Grab the variance components
        var_comp <- as.data.frame(VarCorr(model))
        
        # Get the variance of the breeding lines (Random Effects)
        var_g <- var_comp[1, "vcov"]
        # Grab the total variance of the population
        var_p <- var_comp[2, "vcov"]
        # Calculate R
        r = var_g / (var_g + (var_p / r))
        
        r
}

# Construct a list of all the mixed models of traits setting the sample_ids as random effects and block as a fixed effect
model_list = list(
        'forage_yield_kg_ha' = fy_mm <- lmer(forage_yield_kg_ha ~ 1 + (1|sample_id) + block, data=ibf_10mm_summary),
        'dry_matter_pct' = dm_mm <- lmer(dry_matter_pct ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'moisture_pct' = moi_mm <- lmer(moisture_pct ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'adf_pct' = adf_mm <- lmer(adf_pct ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'ndf_pct' = ndf_mm <- lmer(ndf_pct ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'crude_protein_pct' = cp_mm <- lmer(crude_protein_pct ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'relative_feed_value' = rfv_mm <- lmer(relative_feed_value ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'total_digestible_nutrients' = tdn_mm <- lmer(total_digestible_nutrients ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'neg_adf' = neg_mm <- lmer(neg_adf ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'nel_adf'= nel <- lmer(nel_adf ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'nem_adf' = nem <- lmer(nem_adf ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary)
)

# Extract and store the R values
r_values <- c()
for (name in names(model_list)) {
        print(name)
        model = model_list[[name]]
        r = calc_repeatability(model, r=4)
        r_values <- c(r_values, r)
}

# Build a dataframe with the repeatability values
r_df <- data.frame('test_trait' = names(model_list), 'r' = r_values)

# Get the whole population means
ibf_population_means <- ibf_10mm_summary %>% 
        filter(!sample_id %in% c('HI-GEST 660', 'Kura Clover', 'Sainfoin')) %>% 
        dplyr::select(-c(1:4)) %>% 
        summarize_all(.funs = list('mean' = mean, 'min'=min, 'max'=max, 'var' = var)) %>% 
        pivot_longer(cols = everything(), names_to="test_trait", values_to = "value") %>% 
        separate(test_trait, into=c('test_trait', 'metric'), sep='_(?=[^_]+$)') %>% 
        pivot_wider(names_from = metric)

ibf_population_means <- r_df %>% 
        left_join(ibf_population_means) %>% 
        mutate(range = paste(round(min, 2), round(max, 2), sep = " - ")) %>% 
        dplyr::select(test_trait, mean, range, var, r) 



rownames(ibf_population_means) <- c("Forage Yield kg/Ha", "Dry Matter Pct.", "Moisture Content Pct.", "ADF Pct.", "NDF Pct.", "Crude Protein Pct.", "RFV", "TDN", "NEG ADF", "NEL ADF", "NEM ADF")
ibf_population_means <- ibf_population_means %>% 
        dplyr::select(-test_trait)

kable(
        ibf_population_means, 
        digits = 4, 
        col.names = c('Mean', 'Range', 'Variance', 'R'),
        row.names = TRUE,
        align = 'c'
) %>% 
        kable_material()
```


## Selection Index

Create a selection index such that
$$
I_{penalized} = \bar{I} \cdot(1 - \lambda\cdot\frac{\sigma_I}{\bar{I}})
$$

where $\bar{I}$ is a given replicates selection index calculated
$\sigma_I$ is the standard deviation of the selection indices over all 4 replicates
$\lambda$ is a tuning penalty parameter to downweight samples with a high coefficient of variation


```{r}
# Penalty Term
LAMBDA = 0.5

# Vector of weights for forage_yield, TDN, and crude_protein
W = c(.8, .1, .1)

# Subset for the traits of interest
selection_df <- ibf_10mm_summary %>% 
        dplyr::select(sample_id, forage_yield_kg_ha, total_digestible_nutrients, crude_protein_pct)

selection_ids = selection_df$sample_id

selection_matrix <- as.matrix(selection_df %>% 
        dplyr::select(-sample_id))

SI_vector <- selection_matrix %*% W

SI_df <- data.frame(sample_id = selection_ids, si=SI_vector) %>% 
        group_by(sample_id) %>%
        summarize_all(.funs = list('mean'=mean, 'var'=var)) %>% 
        rowwise() %>% 
        mutate(I_penalize = mean * (1-LAMBDA*(sqrt(var)/mean)))
```


