---
title: "Lumpkin Foundation - Illinois Bundleflower Final Report"
author: "Spencer Barriball, Bo Meyering, Brandon Schlautman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library import, echo=FALSE, results='hide', include=FALSE}
library(tidyverse, warn.conflicts=FALSE, quietly=TRUE)
library(lme4, warn.conflicts=FALSE, quietly=TRUE)
library(sommer, warn.conflicts=FALSE, quietly=TRUE)
library(emmeans, warn.conflicts = FALSE, quietly = TRUE)
library(data.table, warn.conflicts=FALSE, quietly=TRUE)
library(knitr)
library(kableExtra)
library(multcomp)
library(lmerTest)
```

```{r Data Import and Cleaning, include=FALSE, echo=FALSE}
# Read in trial layout and fix sample_name
trial_layout <- fread('../data/trial_layout_metadata.csv') %>% 
        mutate(sample_name = paste('LMP', plots, sep='-')) %>% 
        filter(!trts %in% c('HI-GEST 660', 'Kura Clover', 'Sainfoin')) %>% 
        dplyr::select(-V1) %>% 
        rename("sample_id" = "trts")

# Grab only unique sample_ids
sample_ids <- data.frame(sample_id = unique(trial_layout$sample_id))
# Read in germination data and rearrange
germination_data <- fread('../data/lumpkin_ibf_germination_test.csv') %>% 
        mutate(sample_id = factor(sample_id)) %>% 
        arrange(desc(germination_pct), hard_pct) %>% 
        right_join(sample_ids)

# Read in forage yield data, join to trial layout
yield_data <- fread('../data/lumpkin_ibf_forage_yield.csv') %>% 
        left_join(trial_layout, by='sample_name') %>% 
        dplyr::select(sample_name, sample_id, block, rep, row, column, forage_yield_kg_ha)

# Read in the wet chemistry results
wetchem <- fread('../data/lumpkin_ibf_wet_chemistry_results.csv')

# Pivot wetchem longer
wetchem_longer <- wetchem %>% 
        dplyr::select(-c(moisture_pct, dry_matter_pct)) %>% 
        pivot_longer(cols = c(crude_protein_pct:nel_adf), names_to='test', values_to='results')

# Summarize the yield data
yield_summaries <- yield_data %>% 
        group_by(sample_id) %>% 
        summarize(mean_yield = mean(forage_yield_kg_ha, na.rm=TRUE),
                  sd_yield = sd(forage_yield_kg_ha, na.rm=TRUE),
                  n = n(),
                  se_yield = sd_yield / sqrt(n))

# Pivot germination_data longer
germination_long <- germination_data %>% 
        pivot_longer(cols = c(germination_pct, hard_pct), names_to = 'key', values_to='values')

germination_long$key <- factor(germination_long$key, labels=c('Quick Seed', 'Hard Seed'))

# Read in the model outputs from mlflow runs
model_results = fread("../model_output/runs.csv") %>% 
        mutate(`Test RMSE` = round(`Test RMSE`, 3),
               `Train RMSE` = round(`Train RMSE`, 3)) %>% 
        arrange(`Forage Trait`, desc(`Forage Size`)) %>% 
        dplyr::select(`Forage Trait`, `Forage Size`, `Num. Components`, `Train RMSE`, `Test RMSE`)




```

# Introduction and Summary
## Period Covered by Grant
*Please list the time period this grant covered (e.g. 5/20 - 4/21)*

3/22 - 2/25

## Summary of Work
*Please provide a 3 to 5 sentence summary of your post grant report. Include project name, geographic area served, and project accomplishments. 
Response here.*

[1,000 characters]

### Outcomes
Please indicate which Outcomes you selected on your original application.

* Direct impact on climate, habitat, soil health, or mitigation of extreme weather
* Enhanced public understanding of why your activity is important
* Contribution to climate science that makes future land uses
* Model for communities/organizations how your work improves the lives of people and the environment

### Additional Outcomes
* If previously listed on your application, add your additional outcome not listed above that applies to your request.*

NA

### Metrics
Please indicated which Metrics you selected on your original application.

* Rate of carbon sequestration or expected carbon storage
* Media, community presentations or programs that spread the word on why your project is important
* Volunteers recruited and engaged in your activity
* Acres benefiting from your activity
* ~~Trees newly planted or saved~~
* Scientific papers, presentations, or contacts that add to climate science

### Additional Metrics
*If previously listed on your application, add your additional metric not listed above that applies to your request.*

* Rate of nitrogen export

### Updated Project Evaluation Table
*Please upload your updated Project Evaluation Table to include the Actual data column.  If your actual data missed the target data, please discuss this in your Project Evaluation Narrative.*

Here’s our Project Evaluation Table to be updated.

# Materials and Methods

## Accession Germination Test

## Field Trial Design

## Harvesting

## Sample Processing

## NIR Spectra Sampling
Plant tissues from harvet were ground and filtered using a 1mm sieve .... and scanned using the NeoSpectra NIR handheld scanner ... FILL THIS OUT

## Wet Chemistry
A total of 75 samples out of the entire population were selected for targeted wet chemistry analysis. Samples were selected out of the entire population using the Kennard Stone algorithm to select samples with the greatest Euclidean distance from each other, ensuring that the samples reflected the availble NIR variance in the population. Samples were processed by xyz method, dried and sent to ABC labs for targeted wet chemistry analysis.. FILL THIS OUT


## NIR Modeling
A total of ten wet chemistry metrics were selected for modeling. These included forage moisture content (%), forage dry matter content (%), ADF (units), NDF (units), relative feed value (units), total digestible nutrients (units), net energy gain (Mcal/cwt), net energy maintenance (Mcal/cwt), net energy lactation (Mcal/cwt), and forage yield (kg/Ha) all calculated on a dry matter basis. Spectra were presented as the percent reflectance per wavelength measured. Low quality spectra, i.e. low variance among the response values for all wavelengths, were filtered out of the dataset programatically. All spectra were then linearly resampled by wavelength to ensure equal wavelength intervals between each response value. The spectra were smoothed and detrended using a Savitsky-Golay filter by fitting a $5^{th}$ order polynomial over a window length of 15 frames and taking the second derivative. Standard mean subtraction scaling was applied per wavelength, but were not scaled by the colun variance. Training and testing populations (70% and 30%, respectively) were selected from the calibration samples by using the Kennard-Stone selection algorithm. A partial least squares regression (PLSR) model was fit to the data for each of the target variables, allowing the number of model components to vary from 1 to 16 total components. Root Mean Squared Error (RMSE) defined below was used to evaluate each models performance on the training and testing populations. 
$$
RMSE = \sqrt{\frac{\sum^N_{i=1}(y_i - \hat{y_i})^2}{N}}
$$
where $y_i$ represents the ground truth target value of the $i^{th}$ sample and $\hat{y_i}$ represents the model prediction for that sample. The model with the lowest testing population RMSE was selected as the final model.

## Forage Quality Breeding BLUE and Repeatability
We modeled each of the forage quality trait as $ X $

Repeatabilty was then defined as the variance of the random effect 'Sample ID' divided by the total variance in the model 
$$
\frac{\sigma^2_G}{\sigma^2_G + \frac{\sigma^{2}_{resid.}}{r}}
$$ 
where $r$ is equal to the total number of replicates in the experiment, $\sigma^2_G$ is the sample variance from the random accession effects, and $\sigma^2_{resid.}$ is the residual model variance.

## Selection Index
For each field replicate plot of a given accession, we defined a selection index based on the importance of 3 traits: Forage yield, total digestible nutrients, and crude protein such that the selection index for sample $i$ can be represented as:
$$
SI_i = w \cdot x_i
$$
where $w$ is a column vector of weights that sum to 1, and $x_i$ is a vector of traits from the $i^{th}$ sample expressed as $x_i = [\text{forage yield}, \text{tdn} , \text{crude protein}]$

To balance the tradeoff between the average selection index value and variability within the index estimates for any given accesssion, we found the mean value for each accesssion's samples ($\bar{I} = \sum_{i=1}^{N}SI_i, n=4$) and then penalized it based on the accessions coefficient of variation, resulting in the below definition.

$$
I_{penalized} = \bar{I} \cdot(1 - \lambda\cdot\frac{\sigma_I}{\bar{I}})
$$

where $\bar{I}$ is a given replicates selection index calculated, $\sigma_I$ is the standard deviation of the selection indices over all 4 replicates, $\lambda \in [0,1]$ is a tuning penalty parameter to downweight samples with a high coefficient of variation

# Results
## Germination Results
We performed germination tests on all of the historical accesssions collected. We sowed 100 seeds of each line, and tallied the number of seeds that germinated quickly, and the total number of hard seeds that germinated late. Overall average of quick germination was 45.6% of the seeds, while an additional 18.6% were hard seeds on average for a total mean viability of 64.2%. There was wide variability 



```{r Germination results plot, fig.width=10}
germination_plot <- ggplot(germination_long, aes(x=reorder(sample_id, -total_viable_pct), y=values, fill=key))+
        geom_bar(stat = 'identity', position='stack', width=.9)+
        theme_light()+
        scale_fill_viridis_d(option='D', begin=0.2, end=0.8)+
        labs(x = 'Accession', y='Total Viability (%)', fill="Germination", title="Total Accession Seed Viability")+
        ylim(min=0, max=100)+
        theme(axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              panel.grid.major.x = element_blank(),
              panel.grid.minor.x = element_blank(),
              legend.title = element_blank())

germination_plot
```



## Yield Data
```{r}
trial_layout <- fread('../data/trial_layout_metadata.csv') %>% 
        mutate(sample_name = paste('LMP', plots, sep='-')) %>% 
        dplyr::select(-V1)
yield_data <- fread('../data/lumpkin_ibf_forage_yield.csv') %>% 
        left_join(trial_layout, by='sample_name') %>% 
        dplyr::select(sample_name, trts, block, rep, row, column, forage_yield_kg_ha) %>% 
        rename('sample_id' = 'trts')

# Filter out plots with really low yield
yield_data <- yield_data %>% 
        filter(forage_yield_kg_ha > 500)

yield_summaries <- yield_data %>% 
        group_by(sample_id) %>% 
        summarize(mean_yield = mean(forage_yield_kg_ha, na.rm=TRUE),
                  sd_yield = sd(forage_yield_kg_ha, na.rm=TRUE),
                  n = n(),
                  se_yield = sd_yield / sqrt(n))

yield_mm <- lmer(forage_yield_kg_ha ~ sample_id + (1|block), data=yield_data, )
anova(yield_mm)
# summary(yield_mm)
```


## NIR Modeling

```{r Model Results Import, echo=FALSE}
model_results = fread("../model_output/runs.csv") %>% 
        mutate(`Test RMSE` = round(`Test RMSE`, 3),
               `Train RMSE` = round(`Train RMSE`, 3))

kable(model_results, align = 'c')
```

```{r Yield boxplot, echo=FALSE, include=TRUE}
# Descending yield boxplot
median_values = yield_data %>% 
        group_by(sample_id) %>% 
        summarize(median_yield = median(forage_yield_kg_ha))

yield_data <- yield_data %>% 
        left_join(median_values)


label_df <- yield_data %>% 
        filter(sample_id %in% selections$sample_id) %>% 
        mutate(y_pos = 2000)
yield_plot = ggplot(yield_data, aes(x=reorder(sample_id, -forage_yield_kg_ha, FUN=median), y=forage_yield_kg_ha, fill=median_yield))+
        geom_boxplot()+
        scale_fill_viridis_c(option='D')+
        geom_hline(yintercept=mean(yield_data$forage_yield_kg_ha), linewidth=1, linetype=4)+
        theme_light()+
        labs(x = 'Accession', y='Forage Yield (kg/Ha)', fill = 'Forage Yield')+
        theme(axis.text.x = element_text(angle=90, vjust=0.5),
              axis.ticks.x = element_blank(),
              panel.grid.major.x = element_blank(),
              panel.grid.minor.x = element_blank())+
        geom_text(data=label_df, aes(x = sample_id, y=y_pos), label = "★", color = "red", size = 5)

yield_plot
```

```{r}
# Descending yield boxplot
# median_values = ibf_10mm_summary %>% 
#         left_join(trial_layout %>% dplyr::select(c(sample_name, trts)), by='sample_name') %>% 
#         dplyr::select(-sample_name) %>% 
#         rename('sample_id' = 'trts') %>% 
#         group_by(sample_id) %>% 
#         summarize_all(median) %>% 
#         filter(!sample_id %in% c('HI-GEST 660', 'Kura Clover', 'Sainfoin'))
# 
# ibf_10mm_summary_df = ibf_10mm_summary %>% 
#         left_join(trial_layout %>% dplyr::select(c(sample_name, trts)), by='sample_name') %>% 
#         dplyr::select(-sample_name) %>% 
#         rename('sample_id' = 'trts')
# 
# 
# cp_plot = ggplot(
#         ibf_10mm_summary_df, 
#         aes(x=reorder(sample_id, -adf_pct, FUN=median), y=adf_pct)
#         )+
#         geom_boxplot()+
#         scale_fill_viridis_c(option='D')+
#         theme_light()+
#         labs(x = 'Accession', y='Forage Yield (kg/Ha)', fill = 'Forage Yield')+
#         theme(axis.text.x = element_blank(),
#               axis.ticks.x = element_blank(),
#               panel.grid.major.x = element_blank(),
#               panel.grid.minor.x = element_blank())
# 
# cp_plot
```


```{r}

ggplot(wetchem_longer, aes(x=results, group=test, fill=test))+
        geom_density()+
        facet_wrap(~test, scales = 'free', nrow=2)

crude_protein_plot = ggplot(wetchem, aes(x=reorder(sample_name, -crude_protein_pct, FUN=median), y=crude_protein_pct))+
        geom_boxplot()+
        scale_fill_viridis_c(option='D')+
        # geom_hline(yintercept=mean(yield_data$forage_yield_kg_ha), linewidth=1, linetype=4)+
        theme_light()+
        labs(x = 'Accession', y='Forage Yield (kg/Ha)', fill = 'Forage Yield')+
        theme(axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              panel.grid.major.x = element_blank(),
              panel.grid.minor.x = element_blank())

crude_protein_plot

```

```{r IBF prediction list and summary, echo=FALSE, warning=FALSE, message=FALSE}
ibf_predictions <- list("1mm" = NULL, 
                        "10mm" = NULL)

file_list = dir('../predictions')

for (file_name in file_list) {
        if (startsWith(file_name, 'ibf')) {
                # Extract the forage sample size
                forage_size = str_extract(file_name, "(?<=_)\\d+mm")
                
                # Extract the forage trait
                trait = str_extract(file_name, "(?<=ibf_).*?(?=_\\d+mm)")
                
                file_path = file.path('../predictions', file_name)
                predictions = fread(file_path)

                predictions <- predictions %>% 
                        rename(!!trait := 'value') # Rename the value column dynamically !!(name) evaluates name
                
                if (is.null(ibf_predictions[[forage_size]])) {
                        ibf_predictions[[forage_size]] = predictions
                } else {
                        ibf_predictions[[forage_size]] = ibf_predictions[[forage_size]] %>% 
                                left_join(predictions)
                }
                
                
        }        
}

ibf_1mm_summary <- ibf_predictions[['1mm']] %>% 
        group_by(sample_name) %>% 
        summarize_all(mean) %>% 
        dplyr::select(-rep) %>% 
        left_join(
                trial_layout %>% 
                        dplyr::select(c(sample_name, 1:3)),
                by=c('sample_name')
        ) %>% 
        rename('sample_id' = 'trts') %>% 
        dplyr::select(sample_id, block, plots, sample_name, everything())

ibf_10mm_summary <- ibf_predictions[['10mm']] %>% 
        group_by(sample_name) %>% 
        summarize_all(mean) %>% 
        dplyr::select(-rep) %>% 
        left_join(
                trial_layout %>% 
                        dplyr::select(c(sample_name, 1:3)),
                by=c('sample_name')
        ) %>% 
        rename('sample_id' = 'trts') %>% 
        inner_join(yield_data %>% dplyr::select(sample_name, forage_yield_kg_ha), by='sample_name') %>% 
        dplyr::select(sample_id, block, plots, sample_name, everything())

write_csv(ibf_1mm_summary, 'ibf_1mm_predictions.csv')
write_csv(ibf_10mm_summary, 'ibf_10mm_predictions.csv')


```

## Trait Predictions and Repeatability

```{r repeatability modeling}
# Function to calculate the repeatability of a trait
calc_repeatability <- function(model, r) {
        # Grab the variance components
        var_comp <- as.data.frame(VarCorr(model))
        
        # Get the variance of the breeding lines (Random Effects)
        var_g <- var_comp[1, "vcov"]
        # Grab the total variance of the population
        var_p <- var_comp[2, "vcov"]
        # Calculate R
        r = var_g / (var_g + (var_p / r))
        
        r
}

# Construct a list of all the mixed models of traits setting the sample_ids as random effects and block as a fixed effect
model_list = list(
        'forage_yield_kg_ha' = fy_mm <- lmer(forage_yield_kg_ha ~ 1 + (1|sample_id) + block, data=ibf_10mm_summary),
        'dry_matter_pct' = dm_mm <- lmer(dry_matter_pct ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'moisture_pct' = moi_mm <- lmer(moisture_pct ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'adf_pct' = adf_mm <- lmer(adf_pct ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'ndf_pct' = ndf_mm <- lmer(ndf_pct ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'crude_protein_pct' = cp_mm <- lmer(crude_protein_pct ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'relative_feed_value' = rfv_mm <- lmer(relative_feed_value ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'total_digestible_nutrients' = tdn_mm <- lmer(total_digestible_nutrients ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'neg_adf' = neg_mm <- lmer(neg_adf ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'nel_adf'= nel <- lmer(nel_adf ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary),
        'nem_adf' = nem <- lmer(nem_adf ~ 1 + (1|sample_id) + block, data = ibf_10mm_summary)
)

# Extract and store the R values
r_values <- c()
for (name in names(model_list)) {
        print(name)
        model = model_list[[name]]
        r = calc_repeatability(model, r=4)
        r_values <- c(r_values, r)
}

# Build a dataframe with the repeatability values
r_df <- data.frame('test_trait' = names(model_list), 'r' = r_values)

# Get the whole population means
ibf_population_means <- ibf_10mm_summary %>% 
        filter(!sample_id %in% c('HI-GEST 660', 'Kura Clover', 'Sainfoin')) %>% 
        dplyr::select(-c(1:4)) %>% 
        summarize_all(.funs = list('mean' = mean, 'min'=min, 'max'=max, 'var' = var)) %>% 
        pivot_longer(cols = everything(), names_to="test_trait", values_to = "value") %>% 
        separate(test_trait, into=c('test_trait', 'metric'), sep='_(?=[^_]+$)') %>% 
        pivot_wider(names_from = metric)

ibf_population_means <- r_df %>% 
        left_join(ibf_population_means) %>% 
        mutate(range = paste(round(min, 2), round(max, 2), sep = " - ")) %>% 
        dplyr::select(test_trait, mean, range, var, r) 



rownames(ibf_population_means) <- c("Forage Yield kg/Ha", "Dry Matter Pct.", "Moisture Content Pct.", "ADF Pct.", "NDF Pct.", "Crude Protein Pct.", "RFV", "TDN", "NEG ADF", "NEL ADF", "NEM ADF")
ibf_population_means <- ibf_population_means %>% 
        dplyr::select(-test_trait)

kable(
        ibf_population_means, 
        digits = 4, 
        col.names = c('Mean', 'Range', 'Variance', 'R'),
        row.names = TRUE,
        align = 'c'
) %>% 
        kable_material()
```


## Selection Index
We calculated the penalized selection index for 11 different $\lambda$ weights ranging from 0.0 to 1.0 in 0.1 increment steps, and then calculated the ranking of each accession to see if the rankings were stable (indicating selections that had low variability) or if changing the $\lambda$ parameter caused a lot of rank change in the selection. 

```{r Calculate base selection index, echo=FALSE, include=FALSE}
# Penalty Term
LAMBDA = 0.0

# Vector of weights for forage_yield, TDN, and crude_protein
W = c(.8, .1, .1)

# Subset for the traits of interest
selection_df <- ibf_10mm_summary %>% 
        dplyr::select(sample_id, forage_yield_kg_ha, total_digestible_nutrients, crude_protein_pct)

selection_ids = selection_df$sample_id

selection_matrix <- as.matrix(selection_df %>% 
        dplyr::select(-sample_id))

SI_vector <- selection_matrix %*% W

SI_df <- data.frame(sample_id = selection_ids, si=SI_vector) %>% 
        group_by(sample_id) %>%
        summarize_all(.funs = list('mean'=mean, 'median'=median, 'var'=var)) %>% 
        rowwise() %>% 
        mutate(sd = sqrt(var),
               cv = sd / mean,
               I_bar = mean * (1-LAMBDA*cv)) %>% 
        arrange(desc(I_bar))
```


```{r change of rank}
LAMBDA = c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0)

for(lambda in LAMBDA){
        SI_column_name = paste('SI', lambda, sep='_')
        rank_column_name = paste('rank', lambda, sep = '_')
        
        # Calculate the new SI values
        SI_df[[SI_column_name]] <- SI_df$mean * (1 - lambda * SI_df$cv)
  
        # Rank across all samples (higher SI = better)
        SI_df[[rank_column_name]] <- rank(-SI_df[[SI_column_name]], ties.method = "min")
}

SI_df_ordered <- SI_df %>% 
        arrange(desc(I_bar)) %>% 
        pivot_longer(cols=starts_with('rank'), names_to = 'key', values_to = 'I_penalized')

# Subset for the top 30 lines
subset_df <- SI_df[1:30,]

# Convert to long format for rank columns
df_long <- subset_df %>%
        dplyr::select(sample_id, starts_with("rank_")) %>%
        pivot_longer(cols = starts_with("rank_"),
                     names_to = "lambda",
                     values_to = "rank") %>%
        mutate(lambda = str_remove(lambda, "rank_"),
               lambda = as.numeric(lambda))

labels_df <- df_long %>%
  filter(lambda == 0)

# Plot with labels at lambda = 0
rank_change_plot <- ggplot(df_long, aes(x = lambda, y = rank, group = sample_id, color=rank)) +
        geom_line(linewidth=1) +
        geom_text(data = labels_df,
                  aes(label = sample_id),
                  color = 'black',
                  hjust = .0,  # nudge to the left of the line
                  size = 2.5,
                  nudge_x = -0.1,
                  check_overlap = TRUE) +
        scale_y_reverse() +
        theme_minimal() +
        labs(title = "Rank Changes Across Lambda Values",
             x = expression(lambda),
             y = "Rank (1 = Best)",
             color='Rank') +
        coord_cartesian(clip = "off") +
        theme(panel.grid.minor = element_blank(),
              plot.margin = margin(5.5, 5.5, 5.5, 5.5)) + # extra margin for labels
        scale_color_viridis_c(option='D')+
        scale_x_continuous(limits = c(-.1, 1))

rank_change_plot

```


```{r Change of Rank plot, fig.height=5, fig.width=7}

final_selections <- SI_df %>% 
        arrange(rank_0.5)
final_selections <- selections[1:5,]

selections_summary <- ibf_10mm_summary %>% 
        filter(sample_id %in% selections$sample_id) %>% 
        group_by(sample_id) %>% 
        summarize_at(.vars=vars(adf_pct:forage_yield_kg_ha), .funs = list('mean'=mean, 'sd'=sd)) %>% 
        pivot_longer(cols = adf_pct_mean:forage_yield_kg_ha_sd, names_to = 'key', values_to='values') %>% 
        separate(key, into=c('trait', 'metric'), sep='_(?=[^_]+$)') %>% 
        pivot_wider(names_from='metric', values_from = 'values') %>% 
        mutate(mean_sd = paste(round(mean, 2), "\u00B1", round(sd, 2), sep = " ")) %>% 
        dplyr::select(sample_id, trait, mean_sd) %>% 
        pivot_wider(names_from='trait', values_from = 'mean_sd')

kable(selections_summary) %>% 
        kable_material()
        
```

